Black-swan events in animal populations

### 2016-11-30 text/ms/ms_pnas.tex
Add PNAS ms subfolder

\usepackage{bm}

\input{../analysis/values} % R output

@@

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{
Bayesian $|$
die-offs $|$
ecological surprises $|$
Gompertz $|$
heavy tails $|$
leptokurtic $|$
mass mortality $|$
population dynamics $|$
risk $|$
Student-$t$ distribution
}

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=0.63\textwidth]{../analysis/t-nu-eg2.pdf}
\caption{Illustration of population dynamic models that allow for heavy
tails. (A, B) Probability density for the Student-$t$
distribution with scale parameter of $1$ and different values of \(\nu\). Small
values of \(\nu\) create heavy tails while as \(\nu\) approaches infinity the
distribution approaches the normal distribution. (C to E)
Simulated population dynamics from a Gompertz model with process noise drawn
from Student-$t$ distributions with three values of \(\nu\). Coloured dots in
panels C and D represent jumps with less than a 1 in 1000
chance of occurring in a normal distribution. (F to H)
Estimates of \(\nu\) from models fit to the times series in panels C to
E. Shown are posterior samples (histograms), median and interquartile
range of the posterior (IQR, dots and line segments), and the exponential prior
on \(\nu\) (dashed lines). Colour shading behind panels F to H
illustrates the approximate region of heavy tails.}
\label{fig:1}
\end{figure*}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.28\textwidth]{../analysis/nu-coefs-2.pdf}
\caption{Estimates of population dynamic heavy-tailedness for
populations of birds, mammals, insects, and fishes. (A to D)
Small values of \(\nu\) suggest heavy-tailed black-swan dynamics. Vertical
points and line segments represent posterior medians and 50\% / 90\% credible
intervals for individual populations. Inset plots show probability that \(\nu
< 10\) for populations arranged by taxonomic order and sorted by decreasing
mean Pr(\(\nu < 10\)). Taxonomic orders with three or fewer populations in
panel A are omitted for space. Red to yellow points highlight populations
with a high to moderately high probability of heavy-tailed dynamics.
}
\label{fig:2}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.26\textwidth]{../analysis/order-posteriors-covariates.pdf}
\caption{Standardized probabilities and covariates of heavy-tailed
dynamics. (A) Taxonomic-order-level posterior densities of Pr(\(\nu
< 10\)) after accounting for time-series length in a hierarchical model. Dotted
vertical line in A indicates the median expected Pr(\(\nu < 10\)) from the
prior distribution. Colour shading refers to taxonomic class (yellow: fishes,
green: insects, purple: birds, and red: mammals). Estimates are standardized at
the geometric mean of time-series length across all the data (approximately 27
time steps) and shown for orders with \(\ge 5\) populations. (B)
Posterior densities for potential covariates of Pr(\(\nu < 10\)). In both
panels, short vertical line segments within the density polygons indicate
medians.
}
\label{fig:3}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.45\textwidth]{../analysis/skew-fig.pdf}
\caption{Heavy-tailed process noise tends to be downward skewed and
ignoring this can underestimate risk. (A) Illustrations of Student-$t$
distributions with three levels of skewness (\(\gamma\)) and heavy-tails (\(\nu\)).
(B) Posterior density of the skewness parameters aggregated across
populations grouped into heavy-tailed (\(\hat{\nu} < 10\)), slightly
heavy-tailed (\(10 \leq \hat{\nu} \leq 70\)), and normal-tailed populations
(\(\hat{\nu} > 70\)). Approximate mid-values from B are illustrated in
panel A. (C) Example time series of heavy-tailed populations with
different levels of skewness. Red dots highlight likely heavy-tailed events.
Forecasts (grey regions) show median (solid lines) and lower 99\% credible
intervals (dotted lines) of abundance. Black and red lines indicate forecasts
from Gompertz models with lognormal and skew-$t$ process noise, respectively.
}
\label{fig:4}
\end{figure}

@@

\matmethods{

\subsection{Data}

@@

\subsection{Covariates of heavy-tailed dynamics} We fit a hierarchical beta
regression model to the predicted probability of heavy tails, Pr(\(\nu < 10\)),
to investigate potential covariates of heavy-tailed dynamics. We obtained
maximum lifespan and body-size data from Ref.~\cite{brook2006a}.
Beta regression is useful when response data are on a continuous scale
between zero and one \cite{ferrari2004}. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 10) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + \bm{X_i} \bm{\beta}),
  \: \\
  &\quad \, \text{for } i = 1, \dots, 609\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where \(A\) and \(B\) represent the beta distribution shape parameters;
\(\mu_i\) represents the predicted value for population \(i\), class \(j\),
order \(k\), and species \(l\); \(\phi_\mathrm{disp}\) represents the
dispersion parameter; and \(\bm{X_i}\) represents a vector of predictors (such as
lifespan) for population \(i\) with associated \(\bm{\beta}\) coefficients. The
intercepts are allowed to vary from the overall intercept \(\alpha\) by
taxonomic class (\(\alpha^\mathrm{class}_j\)), taxonomic order
(\(\alpha^\mathrm{order}_k\)), and species (\(\alpha^\mathrm{species}_l\)) with
standard deviations \(\sigma_{\alpha \; \mathrm{class}}\), \(\sigma_{\alpha \;
\mathrm{order}}\), and \(\sigma_{\alpha \; \mathrm{species}}\). Where possible,
we also allowed for error distributions around the predictors by incorporating
the standard deviation of the posterior samples for the Gompertz parameters
\(\lambda\), \(b\), and \(\log \sigma\) around the mean point value as normal
distributions (not shown in the above equation).

@@

% Bibliography
\bibliography{jshort,ms}

>>>>

% \usepackage{bm}

% \input{../analysis/values} % R output
\newcommand{\basePriorMean}{102}
\newcommand{\basePriorMedian}{71}
\newcommand{\basePriorProbHeavy}{7.7}
\newcommand{\medianTimeSteps}{26}
\newcommand{\meanTimeSteps}{30.2}
\newcommand{\minTimeSteps}{20}
\newcommand{\maxTimeSteps}{117}
\newcommand{\birdN}{191}
\newcommand{\insectsN}{182}
\newcommand{\mammalsN}{125}
\newcommand{\fishN}{108}
\newcommand{\birdNH}{14}
\newcommand{\insectsNH}{5}
\newcommand{\mammalsNH}{6}
\newcommand{\fishNH}{0}
\newcommand{\birdPH}{7}
\newcommand{\insectsPH}{3}
\newcommand{\mammalsPH}{5}
\newcommand{\fishPH}{0}
\newcommand{\NOrdersHeavy}{16}
\newcommand{\POrdersHeavy}{41}
\newcommand{\baseFiftyObsFiftySwitch}{8}
\newcommand{\baseSeventyFiveObsFiftySwitch}{2}
\newcommand{\totalHeavyFifty}{26}
\newcommand{\totalHeavySeventyFive}{17}
\newcommand{\baseFiftyObsFiftySwitchPerc}{31}
\newcommand{\baseSeventyFiveObsFiftySwitchPerc}{8}
\newcommand{\baseNuTenObsTenSwitch}{8}
\newcommand{\baseNuTen}{26}
\newcommand{\baseNuFiveObsTenSwitch}{2}
\newcommand{\pHeavyNThirty}{0.13}
\newcommand{\pHeavyNSixty}{0.20}
\newcommand{\pIncHeavyNThirtyNSixty}{1.6}
\newcommand{\obsErrorNuFivePerc}{75}
\newcommand{\modelsNoConvergeAROne}{1}
\newcommand{\modelsNoConvergeAROneHeavyBase}{0}
\newcommand{\percImputedPops}{17}
\newcommand{\percImputedPoints}{0.7}
\newcommand{\nuCoefPopN}{606}
\newcommand{\AvesRangePerc}{4--8}
\newcommand{\InsectaRangePerc}{2--3}
\newcommand{\MammaliaRangePerc}{4--6}
\newcommand{\OsteichthyesRangePerc}{0}
\newcommand{\overallMinPerc}{3}
\newcommand{\overallMaxPerc}{5}
\newcommand{\overallBasePerc}{4}
\newcommand{\NPops}{609}
\newcommand{\NOrders}{39}
\newcommand{\NClasses}{7}
\newcommand{\interpPointsPerc}{1}
\newcommand{\nBSUp}{8}
\newcommand{\nBSDown}{51}
\newcommand{\ratioBSDownToUp}{6.4}
\newcommand{\percBSDown}{86}
\newcommand{\crashUnderRange}{1.1--2}
\newcommand{\crashUnderMedian}{1.3}
\newcommand{\probDensSkewedForHeavyPops}{86}
\newcommand{\percNormPopsNotSkewed}{88}

@@

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{
die-offs $|$
ecological surprises $|$
mass mortality $|$
population dynamics $|$
risk
}

@@

\begin{figure*}[htb]
\centering
\includegraphics[width=0.63\textwidth]{fig1.pdf}
\caption{Illustration of population dynamic models that allow for heavy
tails. (A, B) Probability density for the Student-$t$
distribution with scale parameter of $1$ and different values of \(\nu\). Small
values of \(\nu\) create heavy tails while as \(\nu\) approaches infinity the
distribution approaches the normal distribution. (C to E)
Simulated population dynamics from a Gompertz model with process noise drawn
from Student-$t$ distributions with three values of \(\nu\). Coloured dots in
panels C and D represent jumps with less than a 1 in 1000
chance of occurring in a normal distribution. (F to H)
Estimates of \(\nu\) from models fit to the times series in panels C to
E. Shown are posterior samples (histograms), median and interquartile
range of the posterior (IQR, dots and line segments), and the exponential prior
on \(\nu\) (dashed lines). Colour shading behind panels F to H
illustrates the approximate region of heavy tails.}
\label{fig:1}
\end{figure*}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.28\textwidth]{fig2.pdf}
\caption{Estimates of population dynamic heavy-tailedness for
populations of birds, mammals, insects, and fishes. (A to D)
Small values of \(\nu\) suggest heavy-tailed black-swan dynamics. Vertical
points and line segments represent posterior medians and 50\% / 90\% credible
intervals for individual populations. Inset plots show probability that \(\nu
< 10\) for populations arranged by taxonomic order and sorted by decreasing
mean Pr(\(\nu < 10\)). Taxonomic orders with three or fewer populations in
panel A are omitted for space. Red to yellow points highlight populations
with a high to moderately high probability of heavy-tailed dynamics.
}
\label{fig:2}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.26\textwidth]{fig3.pdf}
\caption{Standardized probabilities and covariates of heavy-tailed
dynamics. (A) Taxonomic-order-level posterior densities of Pr(\(\nu
< 10\)) after accounting for time-series length in a hierarchical model. Dotted
vertical line in A indicates the median expected Pr(\(\nu < 10\)) from the
prior distribution. Colour shading refers to taxonomic class (yellow: fishes,
green: insects, purple: birds, and red: mammals). Estimates are standardized at
the geometric mean of time-series length across all the data (approximately 27
time steps) and shown for orders with \(\ge 5\) populations. (B)
Posterior densities for potential covariates of Pr(\(\nu < 10\)). In both
panels, short vertical line segments within the density polygons indicate
medians.
}
\label{fig:3}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.45\textwidth]{fig4.pdf}
\caption{Heavy-tailed process noise tends to be downward skewed and
ignoring this can underestimate risk. (A) Illustrations of Student-$t$
distributions with three levels of skewness (\(\gamma\)) and heavy-tails (\(\nu\)).
(B) Posterior density of the skewness parameters aggregated across
populations grouped into heavy-tailed (\(\hat{\nu} < 10\)), slightly
heavy-tailed (\(10 \leq \hat{\nu} \leq 70\)), and normal-tailed populations
(\(\hat{\nu} > 70\)). Approximate mid-values from B are illustrated in
panel A. (C) Example time series of heavy-tailed populations with
different levels of skewness. Red dots highlight likely heavy-tailed events.
Forecasts (grey regions) show median (solid lines) and lower 99\% credible
intervals (dotted lines) of abundance. Black and red lines indicate forecasts
from Gompertz models with lognormal and skew-$t$ process noise, respectively.
}
\label{fig:4}
\end{figure}

@@

\matmethods{
\subsection{Data}

@@

\subsection{Covariates of heavy-tailed dynamics} We fit a hierarchical beta
regression model to the predicted probability of heavy tails, Pr(\(\nu < 10\)),
to investigate potential covariates of heavy-tailed dynamics. We obtained
maximum lifespan and body-size data from Ref.~\cite{brook2006a}.
Beta regression is useful when response data are on a continuous scale
between zero and one \cite{ferrari2004}. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 10) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + \pmb{X_i} \pmb{\beta}),
  \: \\
  &\quad \, \text{for } i = 1, \dots, 609\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where \(A\) and \(B\) represent the beta distribution shape parameters;
\(\mu_i\) represents the predicted value for population \(i\), class \(j\),
order \(k\), and species \(l\); \(\phi_\mathrm{disp}\) represents the
dispersion parameter; and \(\pmb{X_i}\) represents a vector of predictors (such as
lifespan) for population \(i\) with associated \(\pmb{\beta}\) coefficients. The
intercepts are allowed to vary from the overall intercept \(\alpha\) by
taxonomic class (\(\alpha^\mathrm{class}_j\)), taxonomic order
(\(\alpha^\mathrm{order}_k\)), and species (\(\alpha^\mathrm{species}_l\)) with
standard deviations \(\sigma_{\alpha \; \mathrm{class}}\), \(\sigma_{\alpha \;
\mathrm{order}}\), and \(\sigma_{\alpha \; \mathrm{species}}\). Where possible,
we also allowed for error distributions around the predictors by incorporating
the standard deviation of the posterior samples for the Gompertz parameters
\(\lambda\), \(b\), and \(\log \sigma\) around the mean point value as normal
distributions (not shown in the above equation).

@@

% Bibliography
% \bibliography{jshort,ms}
\begin{thebibliography}{10}

\bibitem{sornette2002}
Sornette D (2002) Predictability of catastrophic events: Material rupture,
  earthquakes, turbulence, financial crashes, and human birth.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.} 99(suppl 1):2522--2529.

\bibitem{albeverio2006}
Albeverio S, Jentsch V, Kantz H, eds. (2006) {\em Extreme Events in Nature and
  Society}, Frontiers collection.
\newblock (Springer, Berlin).

\bibitem{taleb2007}
Taleb NN (2007) {\em The Black Swan: The Impact of the Highly Improbable}.
\newblock (Random House).

\bibitem{may2008}
May RM, Levin SA, Sugihara G (2008) Complex systems: ecology for bankers.
\newblock {\em Nature} 451(7181):893--895.

\bibitem{sornette2009}
Sornette D (2009) Dragon-kings, black swans and the prediction of crises.
\newblock {\em International Journal of Terraspace Science and Engineering}
  2(1):1--18.

\bibitem{newman2005}
Newman MEJ (2005) Power laws, {Pareto} distributions and {Zipf}'s law.
\newblock {\em Contemporary Physics} 46(5):323--351.

\bibitem{fey2015}
Fey SB et~al. (2015) Recent shifts in the occurrence, cause, and magnitude of
  animal mass mortality events.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.} 112(4):1083--1088.

\bibitem{harnik2012}
Harnik PG et~al. (2012) Extinctions in ancient and modern seas.
\newblock {\em Trends Ecol. Evolut.} 27(11):608--617.

\bibitem{gould1977}
Gould SJ, Eldredge N (1977) Punctuated equilibria: The tempo and mode of
  evolution reconsidered.
\newblock {\em Paleobiology} 3(02):115--151.

\bibitem{mangel1994}
Mangel M, Tier C (1994) Four facts every conservation biologists should know
  about persistence.
\newblock {\em Ecology} 75(3):607--614.

\bibitem{ipcc2012}
{IPCC} (2012) Managing the risks of extreme events and disasters to advance
  climate change adaptation in {\em A Special Report of Working Groups I and II
  of the Intergovernmental Panel on Climate Change}, eds.{} Field CB et~al.
\newblock (Cambridge University Press, Cambridge, UK, and New York, NY, USA).

\bibitem{brook2006a}
Brook BW, Traill LW, Bradshaw CJA (2006) Minimum viable population sizes and
  global extinction risk are unrelated.
\newblock {\em Ecol. Lett.} 9(4):375--382.

\bibitem{knape2012}
Knape J, de~Valpine P (2012) Are patterns of density dependence in the {Global
  Population Dynamics Database} driven by uncertainty about population
  abundance?
\newblock {\em Ecol. Lett.} 15(1):17--23.

\bibitem{keitt1998}
Keitt TH, Stanley HE (1998) Dynamics of {North American} breeding bird
  populations.
\newblock {\em Nature} 393(6682):257--260.

\bibitem{allen2001}
Allen A, Li BL, Charnov E (2001) Population fluctuations, power laws and
  mixtures of lognormal distributions.
\newblock {\em Ecol. Lett.} 4(1):1--3.

\bibitem{halley2002}
Halley J, Inchausti P (2002) Lognormality in ecological time series.
\newblock {\em Oikos} 99(3):518--530.

\bibitem{segura2013}
Segura AM, Calliari D, Fort H, Lan BL (2013) Fat tails in marine microbial
  population fluctuations.
\newblock {\em Oikos} 122(12):1739--1745.

\bibitem{young1994}
Young TP (1994) Natural die-offs of large mammals: implications for
  conservation.
\newblock {\em Conserv. Biol.} 8(2):410--418.

\bibitem{gerber2001}
Gerber LR, Hilborn R (2001) Catastrophic events and recovery from low densities
  in populations of otariids: implications for risk of extinction.
\newblock {\em Mammal. Rev.} 31(2):131--150.

\bibitem{reed2003}
Reed DH, O'Grady JJ, Ballou JD, Frankham R (2003) The frequency and severity of
  catastrophic die-offs in vertebrates.
\newblock {\em Anim. Conserv.} 6(2):109--114.

\bibitem{saucy1994}
Saucy F (1994) Density dependence in time series of the fossorial form of the
  water vole, \textit{Arvicola terrestris}.
\newblock {\em Oikos} 71(3):381--392.

\bibitem{stafford1971}
Stafford J (1971) The heron population of {England} and {Wales}.
\newblock {\em Bird Study} 18(4):218--221.

\bibitem{potts1980}
Potts GR, Coulson JC, Deans IR (1980) Population dynamics and breeding success
  of the shag, \textit{Phalacrocorax aristotelis}, on the {Farne Islands},
  {Northumberland}.
\newblock {\em J. Anim. Ecol.} 49(2):465--484.

\bibitem{gelman2014}
Gelman A et~al. (2014) {\em Bayesian Data Analysis} ed.{} edition T.
\newblock (Chapman {\&} Hall, Boca Raton, FL).

\bibitem{schindler2010}
Schindler DE et~al. (2010) Population diversity and the portfolio effect in an
  exploited species.
\newblock {\em Nature} 465(7298):609--612.

\bibitem{carpenter2006}
Carpenter SR, Brock WA (2006) Rising variance: a leading indicator of
  ecological transition.
\newblock {\em Ecol. Lett.} 9(3):311--318.

\bibitem{nunez2012}
Nu{\~n}ez MA, Logares R (2012) {Black Swans} in ecology and evolution: the
  importance of improbable but highly influential events.
\newblock {\em Ideas in Ecology and Evolution} 5:16--21.

\bibitem{meehl2004}
Meehl GA, Tebaldi C (2004) More intense, more frequent, and longer lasting heat
  waves in the 21st century.
\newblock {\em Science} 305(5686):994--997.

\bibitem{katz2005}
Katz RW, Brush GS, Parlange MB (2005) Statistics of extremes: modelling
  ecological disturbances.
\newblock {\em Ecology} 86(5):1124--1134.

\bibitem{kirby2009}
Kirby RR, Beaugrand G, Lindley JA (2009) Synergistic effects of climate and
  fishing in a marine ecosystem.
\newblock {\em Ecosystems} 12(4):548--561.

\bibitem{denny2009}
Denny MW, Hunt LJH, Miller LP, Harley CDG (2009) On the prediction of extreme
  ecological events.
\newblock {\em Ecol. Monograph.} 79(3):397--421.

\bibitem{gpdd2010}
{NERC Centre for Population Biology, Imperial College} (2010) {\em {The Global
  Population Dynamics Database Version 2}}.

\bibitem{kot1996}
Kot M, Lewis MA, {van den Driessche} P (1996) Dispersal data and the spread of
  invading organisms.
\newblock {\em Ecology} 77(7):2027--2042.

\bibitem{clark1999}
Clark JS, Silman M, Kern R, Macklin E, HilleRisLambers J (1999) Seed dispersal
  near and far: patterns across temperate and tropical forests.
\newblock {\em Ecology} 80(5):1475--1494.

\bibitem{ward2007}
Ward EJ, Hilborn R, Towell RG, Gerber L (2007) A state--space mixture approach
  for estimating catastrophic events in time series data.
\newblock {\em Can. J. Fish. Aquat. Sci.} 64(6):899--910.

\bibitem{nrc2007}
{NRC} (2007) {\em Successful Response Starts with a Map: Improving Geospatial
  Support for Disaster Management}.
\newblock (The National Academies Press, Washington, {DC}).

\bibitem{carpenter2012}
{Carpenter} SR et~al. (2012) General resilience to cope with extreme events.
\newblock {\em Sustainability} 4(12):3248--3259.

\bibitem{caddy1996}
Caddy JF, McGarvey R (1996) Targets or limits for management of fisheries?
\newblock {\em N. Am. J. Fish. Manage.} 16(3):479--487.

\bibitem{doak2008}
Doak DF et~al. (2008) Understanding and predicting ecological dynamics: are
  major surprises inevitable?
\newblock {\em Ecology} 89(4):952--961.

\bibitem{lindenmayer2010}
Lindenmayer DB, Likens GE, Krebs CJ, Hobbs RJ (2010) Improved probability of
  detection of ecological ``surprises''.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.} 107(51):21957--21962.

\bibitem{thompson2013}
Thompson RM, Beardall J, Beringer J, Grace M, Sardina P (2013) Means and
  extremes: building variability into community-level climate change
  experiments.
\newblock {\em Ecol. Lett.} 16(6):799--806.

\bibitem{gaines1993}
Gaines SD, Denny MW (1993) The largest, smallest, highest, lowest, longest, and
  shortest: Extremes in ecology.
\newblock {\em Ecology} 74(6):1677--1692.

\bibitem{fernandez1998}
Fernandez C, Steel MFJ (1998) On {Bayesian} modeling of fat tails and skewness.
\newblock {\em J. Am. Stat. Assoc.} 93(441):359--371.

\bibitem{hoffman2014}
Hoffman MD, Gelman A (2014) The {No-U-Turn Sampler}: adaptively setting path
  lengths in {Hamiltonian Monte Carlo}.
\newblock {\em J. Mach. Learn. Res.} 15:1593--1623.

\bibitem{stan-manual2015}
{Stan Development Team} (2015) {\em Stan Modeling Language User's Guide and
  Reference Manual, Version 2.12.0}.

\bibitem{carpenter2016}
Carpenter B et~al. (2016) Stan: A probabilistic programming language.
\newblock {\em J. Stat. Soft.} In press.

\bibitem{r2015}
{R Core Team} (2016) {\em R: A Language and Environment for Statistical
  Computing} (R Foundation for Statistical Computing, Vienna, Austria).

\bibitem{inchausti2002}
Inchausti P, Halley J (2002) The long-term temporal variability and spectral
  colour of animal populations.
\newblock {\em Evol. Ecol. Res.} 4:1033--1048.

\bibitem{ferrari2004}
Ferrari S, Cribari-Neto F (2004) Beta regression for modelling rates and
  proportions.
\newblock {\em J. Appl. Stat.} 31(7):799--815.

\bibitem{gelman2008c}
Gelman A (2008) Scaling regression inputs by dividing by two standard
  deviations.
\newblock {\em Statist. Med.} 27:2865--2873.

\bibitem{gelman2006c}
Gelman A (2006) Prior distributions on variance parameters in hierarchical
  models.
\newblock {\em Bayesian Analysis} 1(3):515--533.

\bibitem{gelman2008d}
Gelman A, Jakulin A, Pittau MG, Su YS (2008) A weakly informative default prior
  distribution for logistic and other regression models.
\newblock {\em Ann. Appl. Stat.} 2(4):1360--1383.

\bibitem{glmmadmb}
Skaug H, Fournier D, Bolker B, Magnusson A, Nielsen A (2015) {\em {Generalized
  Linear Mixed Models using AD Model Builder}}.
\newblock R package version 0.8.1.

\end{thebibliography}
